Airspace configuration decision based on Monte-Carlo tree search

### 2018-04-18 article/article.tex
(~) moved notations to head
(+) Rewritten a bit algorithms, new notations

\DeclareMathOperator{\argmax}{argmax}

@@

@@

\subsubsection{Extending the search tree}
\paragraph{Selection}
The aim of the selection is to choose from a set of actions, the most promising
one\footnote{i.e.\ possibly leading to the best evaluation} according to a
\emph{tree policy}. This type of problem is addressed by bandits methods where
a multi armed bandit aims to always activate the slot machine giving the highest
reward.

The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
It uses the UCB for Upper Confidence Bound as tree policy (introduced
in~\cite{kocsis2006bandit}). Let \(F\) be the node from which an action must be
selected to go deeper in the tree, \(a\) an action available from \(F\),
\(\mu_{F,a}\) the expected outcome from the node obtained by applying action
\(a\) on \(F\), \(T_F\) the number of simulations carried out from node \(F\)
(which includes any simulation from nodes in any subtree of \(F\)) and
\(T_{F,a}\) the number of simulations carried out from the node resulting from
the action \(a\) applied to node \(F\), \(\beta\) a constant. The selected node
is thus the one maximising
\begin{equation}
  \text{UCB} = \mu_{F,a} + 2 \beta \sqrt{\frac{2 \log T_f}{T_{F,a}}}
\end{equation}

\paragraph{Expansion}
The above step is carried out only if, considering a node \(F\), a simulation
has been run through each child available from \(F\). In the case where one or
more children haven't been explored yet, one of the unexplored children is
chosen and a simulation is started from it.
\begin{algorithm}
  \caption{Extending the search tree}
  \begin{algorithmic}
    \Function{TreePolicy}{$v$}
    \State{} \([v_1, \dots, v_n] \gets\) \Call{Select}{$v$}
    \If{$v_n$ is terminal}
    \State{} \Return{$[v_1, \dots, v_n]$}
    \Else{}
    \State{} \([v_1, \dots, v_{n+1}] \gets\)
    \Call{Expand}{$[v_1, \dots, v_n]$}
    \State{} \Return{$[v_1, \dots, v_{n+1}]$}
    \EndIf{}
    \EndFunction{}
  \end{algorithmic}
  \begin{algorithmic}
    \Function{Select}{$v, a$}
    \If{$v$ is terminal or fully expanded}
    \State{} \Return{$[v, a]$}
    \Else{}
    \State{} fav\(\gets\) \Call{BestChild}{$v$}
    \State{} \Return{\Call{Select}{fav, $[v, a]$}}
    \EndIf{}
    \EndFunction{}
  \end{algorithmic}
  \begin{algorithmic}
    \Function{BestChild}{$v$}
    \State{} \Return{$\argmax\{\Call{UCB}{v'} |
    v' \text{\,children of\,} v\}$}
    \EndFunction{}
  \end{algorithmic}
\end{algorithm}

@@

\section*{Notations}

\begin{itemize}
  \item $m$: elementary module
  \item $S$: sector control position, group of modules
  \item $P$: partition of the airspace, group of sectors
  \item $t$: current time
  \item $C(P, t)$: cost of a given partition $P$ at time $t$
  \item $s$: state, in this article a partition $P$ and a time $t$
  \item $h(s)$: heuristic that gives for a state $s$ an estimation of the cost
  \item $\mu_{F, a}$: mean of all reward comming from node $F$ with action $a$
  \item $T_F$: number of simuation where the final path include the node $F$
  \item $T_{F, a}$: number of simuation where the final path include the
    node $F$ with action $a$
\end{itemize}

>>>>

\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\cons}{::}

@@

\section*{Notations}
\begin{itemize}
  \item $m$: elementary module
  \item $S$: sector control position, group of modules
  \item $P$: partition of the airspace, group of sectors
  \item $t$: current time
  \item $C(P, t)$: cost of a given partition $P$ at time $t$
  \item $s$: state, in this article a partition $P$ and a time $t$
  \item $h(s)$: heuristic that gives for a state $s$ an estimation of the cost
  \item $\mu_{v, a}$: mean of all reward comming from node $v$ with action $a$
  \item $T_v$: number of simulation where the final path include the node $v$
  \item $T_{v, a}$: number of simulation where the final path include the
    node $F$ with action $a$
  \item \(\mathcal{A}(v)\) the available actions from a node \(v\)
  \item \(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\) the
    production function, giving the resulting state of action \(a\) applied on a
    state \(s\).
\end{itemize}

@@

\subsubsection{Extending the search tree}
\paragraph{Selection~\ref{alg:tree_pol}}
The aim of the selection is to choose from a set of actions, the most promising
one\footnote{i.e.\ possibly leading to the best evaluation} according to a
\emph{tree policy}. This type of problem is addressed by bandits methods where
a multi armed bandit aims to always activate the slot machine giving the highest
reward.

The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
It uses the UCB for Upper Confidence Bound as tree policy (introduced
in~\cite{kocsis2006bandit}). Let \(v\) be the node from which an action must be
selected to go deeper in the tree, \(a\) an action available from \(v\),
\(\mu_{v,a}\) the expected outcome from the node obtained by applying action
\(a\) on \(v\), \(T_v\) the number of simulations carried out from node \(v\)
(which includes any simulation from nodes in any subtree of \(v\)) and
\(T_{v,a}\) the number of simulations carried out from the node resulting from
the action \(a\) applied to node \(v\), \(\beta\) a constant. The selected node
is thus the one maximising
\begin{equation}
  \text{UCB} = \mu_{v,a} + 2 \beta \sqrt{\frac{2 \log T_v}{T_{v,a}}}
\end{equation}

\paragraph{Expansion}
The above step is carried out only if, considering a node \(F\), a simulation
has been run through each child available from \(F\). In the case where one or
more children haven't been explored yet, one of the unexplored children is
chosen and a simulation is started from it.
\begin{algorithm}
  \caption{UCT algorithm}\label{alg:tree_pol}
  \begin{algorithmic}
    \Function{TreePolicy}{path of nodes $\pi$}
    \State{} $v \gets$ head of $\pi$
    \State{} \Call{Deploy}{$v$}
    \If{$v$ is terminal}
    \State{}\Return{$\pi$}
    \Else{}
    \If{$v$ is not fully expanded}
    \State{}\Return{\Call{Expand}{$v$} $\cons \pi$}
    \Else{}
    \State $f \gets$ \Call{BestChild}{$v$}
    \State{}\Return{%
      \Call{TreePolicy}{$f \cons \pi$}
    }
    \EndIf{}
    \EndIf{}
    \EndFunction{}
  \end{algorithmic}
  \begin{algorithmic}
    \Function{BestChild}{$v$}
    \State{} \Return{$\argmax\{\Call{UCB}{v'} |
    v' \text{\,children of\,} v\}$}
    \EndFunction{}
  \end{algorithmic}
\end{algorithm}
\begin{algorithm}
  \caption{%
    Expansion (\(p\colon \mathcal{S} \times \mathcal{A} \to \mathcal{S}\)
    is the transition function)
  }
  \begin{algorithmic}
    \Function{Expand}{node $v$}
    \State{} $\ell \gets \{p(v, a) | a\in\mathcal{A}(v), T_{v,a} = 0\}$
    \State{}\Return{random element from $\ell$}
    \EndFunction{}
  \end{algorithmic}
\end{algorithm}

@@
